Limits to Spontaneous Resets in the Double-Drift Illusion

### 2019-12-18 DoubleDriftResetLimits.Rmd
updated notebook with current bounded tracking plots and statistics

---
title: "Limits to Spontaneous Resets in the Double-Drift Illusion"
output: html_notebook
---

@@

```{r}
# The functions here allow downloading the collected data from the OSF project:
source('R/data.R')
# This file has some common helper functions:

# The functions in here are used on the data of the bounded tracking experiments:
source('R/tracking_bounded.R')
# And finally some code for the re-tracing experiment...
source('R/retrace_onePass_V4.R')
# As well as models fit to that data:
source('R/models.R')
```

@@

Here we plot the trajectories recorded during this task (each subplot essentially shows 8 trials overlayed):

```{r fig.width=7, fig.height=6}
# from R/tracking_bounded.R:
plotBoundedTracking()
```

@@

# Experiment 1b: continuous tracking

Therefore, we did another experiment that mimicked the first, but with two changes, all designed to increase the occurrence of resets. First, we only used -3 and 3 cycles per second for the internal motion as this would increase the strength of the illusion and perhaps lead to a higher chance of reset. Second, the external motion was halved as earlier work showed that resets often occur after around 2 seconds [@Lisi2015 ??], and 2 seconds was how long one pass (from one end of the screen to the other) took in the first experiment. So now, the double-drift stimulus would do a single pass in 4 seconds instead of 2 seconds. This allows more opportunity for resets, but might also make it easier to track the gabor's position, so that errors at the direction changes in the far and close end of the workspace are less likely.

>>>>

---
title: "Limits to Spontaneous Resets in the Double-Drift Illusion"
output:
  word_document: default
  html_notebook: default
---

@@

```{r}
# The functions here allow downloading the collected data from the OSF project:
source('R/data.R')
# This file has some common helper functions:
source('R/common.R')
# The functions in here are used on the data of the bounded tracking experiments:
source('R/tracking_bounded.R')
# And finally some code for the re-tracing experiment...
source('R/retrace_onePass_V4.R')
# As well as models fit to that data:
source('R/models.R')
```

@@

First, we process the raw data into a more useful format for our analyses:

```{r}
# from R/tracking_bounded.R:
getBoundedTrackingStandardizedSegments(segmentpoints=101, version=1)
getSegmentDirections(version=1)
```

Then we plot the trajectories recorded during this task (each subplot essentially shows 8 trials overlayed), and in the top row we show the average heading as a thicker line, as well as a 2D polar histogram showing the distribution of headings.

```{r fig.width=7, fig.height=6, echo=F}
# from R/tracking_bounded.R:
plotBoundedTracking()
```

@@

# Experiment 1b: continuous tracking with 4 second segments

Therefore, we did another experiment that mimicked the first, but with two changes, all designed to increase the occurrence of resets. First, we only used -3 and 3 cycles per second for the internal motion as this would increase the strength of the illusion and perhaps lead to a higher chance of reset. Second, the external motion was halved as earlier work showed that resets often occur after around 2 seconds [@Lisi2015 ??], and 2 seconds was how long one pass (from one end of the screen to the other) took in the first experiment. So now, the double-drift stimulus would do a single pass in 4 seconds instead of 2 seconds. This allows more opportunity for resets, but might also make it easier to track the gabor's position, so that errors at the direction changes in the far and close end of the workspace are less likely. Since the passes are longer, we use only 3 full passes in each trial, but run 32 trials in total (instead of the 8 per condition).

Before calculating heading, we've normalized the data, so that the direction of all segments _should_ be the same.

First, here is a figure with the heading distributions across time for all participants as well as the group average:

```{r fig.width=8, fig.height=4, echo=F}
plotBoundedTracking_V2(target='inline')
```

Again, we see no time-dependent decrease of heading towards zero, except perhaps at the start of the data for participant 12. Since these are preceded by very large headings, and then settle into the normal range, this participant might have had more trouble responding to the direction reversals of the gabor. The group average reflects this tendency of participant 12 somewhat, but overall the average heading is very close to what was seen in the first trakcing experiment: 28.9 compared to 27.4 or 26.7 degrees.

# Experiment 2: one-pass tracking and psychophysics

Since it seemed that no clear spontaneous resets occurred in the continuous tracking paradigm, we used a re-tracing approach (c.f. @Nakayama2019).
